---
title: 'Ass 2'
author: "Daniel Li"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, fig.align='center'}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

## a)

Given a match, the probability of an expert failing to identify a match is 1-0.9212 = 0.0788

## b)

Given a match, the probability of a novice failing to identify a match is 1 - 0.7455 = 0.2545

## c)

Using Bayes' theorem, we can conclude the participant is more likely to be a novice. P(expert|fail) = ((1-.92) * .92)/P(fail), P(novice) = ((1-.75) *.75)/fail.
Even without solving for P(fail), we can see that P(expert|fail) is lower.

# Question 2

## a)

P(Positive|Doping) = ((50/59)*(59/1000)/.1 = 0.5 or 50 percent

## b)

P(Negative|Nonuser) = 891/941 = .94685 or about 95 percent

## c)

P(Doping|Positive) = (P(P|D) * P(P))/(P(P|D)* P(P) + P(P|Non) * P(Non)) = (0.5* 0.1)/((0.5 * 0.1) + 0.01 * .9) = 0.8475 or about 85 percent.



# Question 7

## a)

.09 + 0.3 + 0.37 + 0.2 + 0.04 = 1 

## b) 

$P(3 U 4)$ = 0.2 + 0.4 = 0.24

## c 

$P(0 U 1)$ = 0.09 + 0.3 = 0.39


# Question 8

## a)

We can sum up p(y) to make sure it goes to 1

```{r}
.17 + .10 + .11 + .11 + .1 + .1 + 0.07 + 0.05 + 0.03 + 0.02 + 0.02 + 0.02 + 0.02 + 0.02 + 0.01 + 0.01 + 0.01 + 0.01 + 0.01 + 0.005 + 0.005
```
p(y) is also defined for all y

## b)

P(y >= 10) = 0.02 + 0.02 + 0.02 + 0.02 + 0.05 + 0.01 = 0.14

## c) 

mean = $\mu = \Sigma yp(y)$ = 1* .10 + 2 * .11 +  3 * .11 + 4 *  .10 + 5 * .10 + 6 * .07 + 7 * .05 + 8 * .03 + 9 * .02 + 10 * .02 + 11 * .02 +  12 *  .02 + 13 * .02 + 14 * .01 + 15 * .01 + 16 * .01 + 
17 * .01 + 18 * .01 + 19 * .005 + 20 .005

= 4.475

variance = $\sigma^2 = E(y-\mu)^2p(y) $ =  19.608616

## d)

An interval that wil contain at least .75 of the probability is [1,20]

# Question 9

## a)

```{r}
dbinom(10, size = 25, prob = 0.7)
```

## b)

```{r}
pbinom(c(5), size = 25, prob = 0.7)
```

## c)

mean = 25 * .7 = 17.75

sd = sqrt(17.75*(1-.7)) = 2.3

## d) 

It means that, on average, there are about 17.75 foreign nationals that have a PhD in engineering in 25 PhDs, and that in most trials, that number will usually be higher or lower by 2.3.

## Question 10

## a)

```{r}

dmultinom(rep(5, 10), size = 50, rep(1/10,10))

```
## b)

```{r}
dbinom(0, size = 50, prob = 1/10) + dbinom(1, size = 50, prob = 1/10)
```

