---
title: "ASS1"
output: 
  html_document:
    toc: yes
    toc_float: yes
  
date: "2024-02-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Question 1

Grading scale: Grade consists of labs, in class quizzes, exams, chapter quizzes, projects, assignments, and the final exam.

A is 90s and up. B is 80s. C is 60s and 70s. D is 50s and 60s. F is anything lower.

# Question 2

## 2a)

```{r}
ddt <- read.csv("DDT.csv")

m=with(ddt, as.numeric(factor(MILE)))
mcol=with(ddt,ifelse(MILE > 3, "green",
                        ifelse(MILE < 3,"blue","green")))
with(ddt,coplot(LENGTH~WEIGHT | (RIVER*SPECIES),pch = 21, col = MILE))
```

## 2b)
Given a river FCM, LCM, and SCM, how will Length increase as a function of Weight?

## 2c)
It stores the MILE data from ddt into the variable m as a numeric type.

## 2d)

Finds how many unique values of m there are.

## 2e)

There are no LMBASS or SMBUFFALO in FCM, LCM, and SCM Rivers.

## 2f)

```{r}
mean(ddt$SPECIES == "CCATFISH" & ddt$RIVER == "FCM")
```
The mean value of ddt found in the sample of CCATFISH in FCM river is  0.04166667


# Question 3

## 3a)

Quantitative

## 3b) 

Quantitative

## 3c) 

Qualitative

## 3d)

Quantitative

## 3e) 

Qualitative

## 3f)

Quantitative

## 3g)

Qualitative

# Question 4

## 4a) 

The four sampling designs are simple random sampling, stratified random sampling, cluster sampling, and systematic sampling.

## 4b)

Simple random sampling: Every individual data point of a set of data is randomly chosen, all with equal chance.

Stratified random sampling: Stratify the data (separate data set into relevant subsets e.g by speed limit on a road) and then randomly sample.

Cluster sampling: Sample multiple, related subsets of data rather than single points. For example, When examining average grocery prices of fruits, group brands of the same fruit into one sample.

Systematic sampling: Systematically select samples by a given metric e.g every 3rd or every 5th point.

# Question 5

## 5a) i)

```{r}
mtbe <- read.csv("MTBE.csv")

ind = sample(1:223,5, replace = FALSE)
mtbe[ind,]
mtbeo=na.omit(mtbe)
```
## 5a) ii) 
```{r}
depth=mtbeo[mtbeo$Aquifier=="Bedrock",]$Depth
sd(depth)
```
# Question 6
## 6a) i)

```{r}
library(readxl)
eq <- read_excel("EARTHQUAKE.xls")
ind = sample(1:2929,30, replace = FALSE)
eq[ind,]
plot(ts(eq$MAGNITUDE))
with(eq,median(MAGNITUDE))
```
# Question 7

## 7a)

They used stratified random sampling to select samples from the rivers.

## 7b)

Fish in the Alabama river.

## 7c)

River, Species

# Question 8 

## 8a) 

Bar graph 

## 8b) 

Number of robots 

## 8c) 

Legs ONLY

## 8d) 




