---
title: "ASS1"
output: 
  html_document:
    toc: yes
    toc_float: yes
  
date: "2024-02-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Number of Questions completed: 14/15
# Question 1

Grading scale: Grade consists of labs, in class quizzes, exams, chapter quizzes, projects, assignments, and the final exam.

A is 90s and up. B is 80s. C is 60s and 70s. D is 50s and 60s. F is anything lower.

# Question 2

## 2a)

```{r}
ddt <- read.csv("DDT.csv")

m=with(ddt, as.numeric(factor(MILE)))
mcol=with(ddt,ifelse(MILE > 3, "green",
                        ifelse(MILE < 3,"blue","green")))
with(ddt,coplot(LENGTH~WEIGHT | (RIVER*SPECIES),pch = 21, col = MILE))
```

## 2b)
Given a river FCM, LCM, and SCM, how will Length increase as a function of Weight?

## 2c)
It stores the MILE data from ddt into the variable m as a numeric type.

## 2d)

Finds how many unique values of m there are.

## 2e)

There are no LMBASS or SMBUFFALO in FCM, LCM, and SCM Rivers.

## 2f)

```{r}
mean(ddt$SPECIES == "CCATFISH" & ddt$RIVER == "FCM")
```
The mean value of ddt found in the sample of CCATFISH in FCM river is  0.04166667


# Question 3

## 3a)

Quantitative

## 3b) 

Quantitative

## 3c) 

Qualitative

## 3d)

Quantitative

## 3e) 

Qualitative

## 3f)

Quantitative

## 3g)

Qualitative

# Question 4

## 4a) 

The four sampling designs are simple random sampling, stratified random sampling, cluster sampling, and systematic sampling.

## 4b)

Simple random sampling: Every individual data point of a set of data is randomly chosen, all with equal chance.

Stratified random sampling: Stratify the data (separate data set into relevant subsets e.g by speed limit on a road) and then randomly sample.

Cluster sampling: Sample multiple, related subsets of data rather than single points. For example, When examining average grocery prices of fruits, group brands of the same fruit into one sample.

Systematic sampling: Systematically select samples by a given metric e.g every 3rd or every 5th point.

# Question 5

## 5a) i)

```{r}
mtbe <- read.csv("MTBE.csv")

ind = sample(1:223,5, replace = FALSE)
mtbe[ind,]
mtbeo=na.omit(mtbe)
```
## 5a) ii) 
```{r}
depth=mtbeo[mtbeo$Aquifier=="Bedrock",]$Depth
sd(depth)
```
# Question 6
## 6a) i)

```{r}
library(readxl)
eq <- read_excel("EARTHQUAKE.xls")
ind = sample(1:2929,30, replace = FALSE)
eq[ind,]
plot(ts(eq$MAGNITUDE))
with(eq,median(MAGNITUDE))
```
# Question 7

## 7a)

They used stratified random sampling to select samples from the rivers.

## 7b)

Fish in the Alabama river.

## 7c)

River, Species

# Question 8 

## 8a) 

Bar graph 

## 8b) 

Number of robots 

## 8c) 

Legs ONLY

## 8d) 

None: 0.1415
Both: 0.0754717
Legs ONLY: 0.5943396
Wheels ONLY:0.1886792

## 8e)


```{r}
pareto <- function(x, mn = "Pareto barplot", ...) {  # x is a vector
  x.tab = table(x)
  xx.tab = sort(x.tab, decreasing = TRUE, index.return = FALSE)
  cumsum(as.vector(xx.tab)) -> cs
  length(x.tab) -> lenx
  bp <- barplot(xx.tab, ylim = c(0,max(cs)),las = 2)
  lb <- seq(0,cs[lenx], l = 11)
  axis(side = 4, at = lb, labels = paste(seq(0, 100, length = 11), "%", sep = ""), las = 1, line = -1, col = "Blue", col.axis = "Red")
  for(i in 1:(lenx-1)){
    segments(bp[i], cs[i], bp[i+1], cs[i+1], col = i, lwd = 2)
  }
  title(main = mn, ...)
}
dat <- rep(c("None", "Both", "Legs ONLY", "Wheels ONLY"),c(15,8,63,20) )
pareto(dat, mn ="Relative frequencies of Robotic Limbs")
```

# Question 9

## 9a)

```{r}
pie(c(6,8,22,3,11), labels = c("Denial of service", "Information Disclosure", "Remote Code 
                               Execution", "Spoofing","Privilege elevation"), main = "Security
    Issues in 2012")
```
Spoofing had the lowest proportion of security issues in 2012.

## 9b)

```{r}
sect <- c("Denial of service", "Information Disclosure", "Remote Code 
                              Execution", "Spoofing", "Privilege Elevation")
sec = c(6,8,22,3,11)

dat <- rep(sect,sec)
pareto(dat, mn = "Securtiy issues in 2012")
```

# Question 10

```{r}
library(plotrix)
swd <- read_excel("SWDEFECTS.xls")
tab=table(swd$defect)
rtab=tab/sum(tab)
round(rtab,2)
pie3D(rtab,labels=list("OK","Defective"),main="pie plot of SWD")
```
# Question 11

## 11a)

```{r}
voltage <- read_excel("VOLTAGE.xls")
head(voltage)
```

# Question 12

```{r}
prob <- c(1.72,2.50 ,2.16, 2.13, 1.06, 2.24, 2.31, 2.03, 1.09, 1.40,
2.57, 2.64, 1.26, 2.05, 1.19, 2.13, 1.27, 1.51, 2.41, 1.95)
quantile(prob, 0.025)
quantile(prob, 0.975)

```

An interval in which 95% of the data will lie between is (1.07425,2.60675)


# Question 13

## 13a)

```{r}
mean(c(3,3,52,7,5,49,5,4,4,5,4))
median(c(3,3,52,7,5,49,5,4,4,5,4))
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
getmode(c(3,3,52,7,5,49,5,4,4,5,4))
```
mean:12.81818
median:5
mode: 5

## 13b)

While I would usually recommend mean, I believe that the median is a better fit here. The 52 and 49 are extreme outliers compared to the large majority.

## 13c)

```{r}
mean(c(40,52,40,43,27))
median(c(40,52,40,43,27))
getmode(c(40,52,40,43,27))
```

Mean:40.4
Median:40
Mode:40

## 13d)
```{r}
mean(c(40,52,40,43,27))
median(c(40,52,40,43,27))
getmode(c(30,16,30,56,22,14))
```
Mean:40.4
Median:40
Mode:30

## 13e)

Since both the medians and means are the same, the total plant coverage distribution is the same on a curve, but on a barplot, it would not be the same.

# Question 14 

# 14a)

```{r}
vec <- c(22922, 20210, 21911, 19225, 18792, 21993, 23059,
20785 ,22781, 23303, 22192, 19462 ,19057, 23017,
20186, 23292, 19408, 24909, 19866, 22891, 23121,
19673 ,23261, 22796, 22355, 19807, 23432, 22625,
22744 ,22426, 19111, 18933, 22417, 19595, 23408,
22809, 19619 ,22738, 18499, 19130, 23220, 22647,
22718, 22779, 19026, 22513, 19740, 22682, 19179,
19404, 22193)
plot(density(vec))
```

## 14b)

Yes, it does seem to be supported. We have 2 different local maxima that have a significant difference.

## 14c) 
```{r}
gala <- vec[vec <= 21000]
galb <- vec[vec > 21000]
mean(gala)
mean(galb)
sd(gala)
sd(galb)
```   
A1775a mean and standard devation: 19462.24,532.2868.
A1775b's mean and standard devation: 22838.47,560.9767.

## 14d)

It would be more likely to be in A1775A, since it falls within the range of local min and max of A1775A's cluster.

# Question 15

```{r}
library(ggplot2)
library(dplyr)
ddt %>% filter(RIVER == "FCM" | RIVER == "LCM" | RIVER == "SCM" | RIVER == "TRM")  %>% ggplot(aes(x=RIVER ,y=LENGTH,fill= SPECIES)) + geom_boxplot() + ggtitle("Daniel Li")
```

